---
title: "2024 NAEP Advanced Performance Dashboard"
toc: true
toc-location: right
format: html
---

The maps below highlight the percent of students scoring advanced on math and reading in 4th and 8th grade by state. Each map is separated into three categories with the states having the least percentage of students scoring advanced highlighted in red, the middle amount in yellow, and the most in green.

## All Students

<iframe src="https://thm43608.github.io/naep24/RYG_advanced_all_math_grade4_map_leaflet.html" width="100%" height="600px" style="border: none;">

</iframe>

<iframe src="https://thm43608.github.io/naep24/RYG_advanced_all_math_grade8_map_leaflet.html" width="100%" height="600px" style="border: none;">

</iframe>

<iframe src="https://thm43608.github.io/naep24/RYG_advanced_all_reading_grade4_map_leaflet.html" width="100%" height="600px" style="border: none;">

</iframe>

<iframe src="https://thm43608.github.io/naep24/RYG_advanced_all_reading_grade8_map_leaflet.html" width="100%" height="600px" style="border: none;">

</iframe>

## Economically Disadvantaged Students

The maps below look at performance for students based on whether the students are economically disadvantaged or not.

<iframe src="RYG_advanced_SESdisadvantaged_math_grade4_map_leaflet" width="100%" height="600px" style="border: none;">

</iframe>

<iframe src="https://thm43608.github.io/naep24/RYG_advanced_SES_not_disadvantaged_math_grade4_map_leaflet.html" width="100%" height="600px" style="border: none;">

</iframe>

You can see the longitudinal changes by economically disadvantaged status in the chart below.

```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readxl)
library(here)
library(here)
df_ed_m4 <- read_excel("../../data/cleaned/Longitudinal/cleaned_2024_math_grade4_economically_disadvantaged_multiyear.Xls")

# Prepare the dataset
df_ed_national <- df_ed_m4 %>%
  filter(Jurisdiction == "National public") %>%
  filter(`Economically disadvantaged status` %in% c("Economically disadvantaged", "Not economically disadvantaged")) %>%
  mutate(`at Advanced` = as.numeric(`at Advanced`)) %>%
  arrange(Year)

df_race_m4 <- read_excel("../../data/cleaned/Longitudinal/cleaned_2024_math_grade4_race_multiyear.Xls")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=5}
ggplot(df_ed_national, aes(x = Year, y = `at Advanced`, color = `Economically disadvantaged status`, group = `Economically disadvantaged status`)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(title = "Trend of 'At Advanced' Performance Over Years",
       subtitle = "Comparison: Economically Disadvantaged vs. Not Economically Disadvantaged",
       x = "Year",
       y = "Percentage of Students at Advanced Level",
       color = "Economic Status") +
  theme_minimal() +
  scale_x_continuous(breaks = unique(df_ed_national$Year)) +
  scale_y_continuous(limits = c(0, max(df_ed_national$`at Advanced`, na.rm = TRUE) + 2))
```

## Difference in Advanced Performance by Race

The visualizations below look at performance for students based on race.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=5}

ggplot(df_race_m4, aes(x = Year, y = `at Advanced`, color = `Race/ethnicity used to report trends, school-reported`, group = `Race/ethnicity used to report trends, school-reported`)) +
  geom_line(linewidth = 1) +         # Line plot for trends
  geom_point(size = 2) +             # Highlight data points
  labs(title = "Trend of 'At Advanced' Performance Over Years",
       subtitle = "Comparison of Race for Math Grade 4",
       x = "Year",
       y = "Percentage of Students at Advanced Level",
       color = "Race") +
  theme_minimal() +
  scale_x_continuous(breaks = unique(df_race_m4$Year)) +  # Ensure all years are displayed
  scale_y_continuous(limits = c(0, max(df_race_m4$`at Advanced`, na.rm = TRUE) + 2))  # Ensure proper y-axis scaling
```

## Performance by Gender for Each State

You can see the math performance by gender for each state for Grades 4 and 8 by [clicking here](https://thm43608.github.io/naep24/math_gender.html){target="_blank"}.

## About the Data and Maps

These maps were created using [NAEP data explorer](https://www.nationsreportcard.gov/ndecore/landing){target="_blank"}.

Tyler Clark created the maps on January 31, 2025. You may contact Tyler by email at [thomas.clark\@wku.edu](thomas.clark@wku.edu).
